## Violin Plots {#violin}

In this section, we introduce how to combine music with R via creating **violin plots**. In fact, the violin plot elegantly combines density plot (Section \@ref(density-plot)) and boxplot (Section \@ref(boxplots)) and into a single plot. 

### The basic violin plot

Let's say we want to generate a basic violin plot for the variable `cholesterol` in the `gm2004` dataset.

```{r, warnings = FALSE}
library(r02pro)
library(ggplot2)
ggplot(data = gm2004, 
       aes(x = "", y = cholesterol)) + 
  geom_violin()
```

To introduce the detail of the generation process of violin plot, it is helpful to review the density plot.
```{r, warnings = FALSE}
ggplot(data = gm2004, 
       aes(y = cholesterol)) +
  geom_density()
```

Looking at these two plots, it is easy to see that the basic violin plot is nothing but a **Mirrored Density Plot** with the kernel density estimates on each side. 

### Violin plot with boxplot

Usually, the violin plot includes the boxplot inside it, providing extra information about the data. To do this, we just add the boxplot layer on top of the violin plot.

```{r, warnings = FALSE}
ggplot(data = gm2004, 
       aes(x = "", y = cholesterol)) + 
  geom_violin() +
  geom_boxplot(width = 0.1)
```

Here, we set the aesthetic `width = 0.1` in the boxplot to make it thinner. 

Just like in the boxplot, we can compare the distributions of a continuous variable for different values of a discrete variable. We can achieve this by mapping the discrete variable to the x axis.

```{r}
ggplot(data = gm2004, 
       aes(x = continent, 
           y = cholesterol)) + 
  geom_violin() +
  geom_boxplot(width = 0.1)
```

We can restrict the x-axis to a subset of the possible `continent` values. 
```{r}
ggplot(data = gm2004, 
       aes(x = continent, 
           y = cholesterol, 
           color = continent)) + 
  geom_violin() +
  geom_boxplot(width = 0.1) + 
  scale_x_discrete(limits=c("Africa", "Americas", "Asia"))
```


Similarly, we can map a third variable to an aesthetic.

```{r}
ggplot(data = gm2004, 
       aes(x = continent, 
           y = cholesterol, 
           fill = gender)) + 
  geom_violin() +
  geom_boxplot(width = 0.1) + 
  scale_x_discrete(limits=c("Africa", "Americas", "Asia"))
```

As you can see, the boxplot doesn't align well inside the violin plot. To fix this issue, you can add the global aesthetic `position = position_dodge(0.9)` to both geoms. 

```{r}
ggplot(data = gm2004, 
       aes(x = continent, 
           y = cholesterol, 
           fill = gender)) + 
  geom_violin(position = position_dodge(0.9)) +
  geom_boxplot(width = 0.1, position = position_dodge(0.9)) + 
  scale_x_discrete(limits=c("Africa", "Americas", "Asia"))
```

You can also try to add other Constant-Valued Aesthetics to both geoms to change their appearances. 

```{r}
ggplot(data = gm2004,
       aes(x = continent,
           y = cholesterol,
           color = continent)) +
  geom_violin(color = "violet",
              size = 2) +
  geom_boxplot(width = 0.1,
               color = "blue",
               size = 1) +
  scale_x_discrete(limits=c("Africa", "Americas", "Asia"))
```

In this section, we introduced the violin plot, which combines the density plot and boxplot into a single visualization. By layering `geom_violin()` and `geom_boxplot()`, you can simultaneously see the distribution shape and summary statistics. When comparing groups with a third variable, remember to use `position = position_dodge(0.9)` to properly align the plots.

### Exercises

1. Using the `sahp` dataset from the **r02pro** package, create a violin plot for `sale_price` grouped by `house_style`, restricting to `"1Story"` and `"2Story"`. Add a boxplot inside with `width = 0.1`.

2. Using the `gm2004` dataset, create a violin plot for `BMI` grouped by `continent`. Map `continent` to the `fill` aesthetic and add a boxplot inside with `width = 0.1` and `position = position_dodge(0.9)`.

3. Modify your plot from Exercise 2 to only include `"Europe"` and `"Asia"`, and further group by `gender` using the `fill` aesthetic. Make sure both the violin and boxplot use `position = position_dodge(0.9)`.

4. Create a violin plot for `cholesterol` in the `gm2004` dataset. Customize the violin with `color = "darkgreen"` and `size = 1.5`, and add a boxplot with `color = "red"` and `width = 0.1`.

