## Interactive Plot with `plotly()` {#plotly}

Until now, all the plots we created are static. In this section, we will introduce a powerful tool named `plotly()` that can make interactive plots. If you haven't done so, you need to first install the R package **plotly**.

```{r, eval = FALSE}
install.packages("plotly")
```

### Converting ggplot2 plots with `ggplotly()`

The easiest way to create interactive plots is to convert an existing **ggplot2** plot using `ggplotly()`. Let's start with a static scatterplot and make it interactive.

```{r}
library(r02pro)
library(ggplot2)
library(plotly)
my_plot <- ggplot(data = gm2004) +
    geom_point(mapping = aes(x = sugar,
                             y = cholesterol))
ggplotly(my_plot)
```

Upon a first look, this may look identical to a regular scatterplot. However, you can try to move your cursor to the points which will show the corresponding coordinates. Some other features offered are available via a bar of buttons on the top right of the plot. Some useful features include

- Download plot as png.
- Zoom: Zoom a region of the plot.
- Pan: Move the plot around.

### Interactive plots with aesthetics

In addition to the vanilla scatterplots, you can use plotly with more complicated plots that involve aesthetics.

```{r}
my_plot_continent <- ggplot(data = gm2004) +
    geom_point(mapping = aes(x = sugar,
                             y = cholesterol,
                             color = continent))
ggplotly(my_plot_continent)
```

Now, you can easily see the `continent` in addition to the `sugar` and `cholesterol` values for each data point.

### Customizing hover text with `tooltip`

By default, `ggplotly()` displays all mapped aesthetics when you hover over a point. You can control which aesthetics appear in the tooltip using the `tooltip` argument.

```{r}
my_plot_tooltip <- ggplot(data = gm2004) +
    geom_point(mapping = aes(x = sugar,
                             y = cholesterol,
                             color = continent,
                             text = country))
ggplotly(my_plot_tooltip, tooltip = c("text", "x", "y"))
```

Here, we mapped the `country` variable to the `text` aesthetic and told `ggplotly()` to only show `text`, `x`, and `y` in the tooltip. This is useful when you want to display additional information beyond the x and y coordinates.

### Interactive boxplots

Let's try to use plotly with some other types of plots.

```{r}
my_box_plot <- ggplot(data = na.omit(sahp),
            aes(x = house_style,
                y = sale_price)) +
  scale_x_discrete(limits=c("1Story", "2Story")) + geom_boxplot(mapping = aes(color = oa_qual > 5))
ggplotly(my_box_plot)
```

For boxplots, we can easily see the values of summary statistics. You are welcome to try out plotly on other types of plots.

### Saving interactive plots

Interactive plots created with **plotly** are HTML widgets. You can save them as standalone HTML files using the `htmlwidgets::saveWidget()` function.

```{r, eval = FALSE}
library(htmlwidgets)
my_interactive <- ggplotly(my_plot)
saveWidget(my_interactive, "my_interactive_plot.html")
```

The saved HTML file can be opened in any web browser and shared with others — no R installation required.

::: {.infobox .caution data-latex="{caution}"}
Interactive plots created with **plotly** only work in HTML output (e.g., the gitbook format). They will not render in PDF or other static formats. If you need to include plots in a PDF document, use static **ggplot2** plots instead.
:::

In this section, we introduced the **plotly** package for creating interactive plots. The `ggplotly()` function provides a convenient way to convert any static **ggplot2** plot into an interactive one, enabling features like hovering to see data values, zooming, and downloading as PNG.

### Exercises

1. Using the `sahp` dataset from the **r02pro** package, create an interactive scatterplot of `liv_area` (x-axis) vs. `sale_price` (y-axis) using `ggplotly()`.

2. Enhance the plot from Exercise 1 by mapping `house_style` to the `color` aesthetic. Hover over points to verify that the house style information is displayed.

3. Create an interactive scatterplot of `sugar` vs. `cholesterol` from the `gm2004` dataset, and customize the tooltip to display the `country` name along with the x and y values.

4. Create an interactive violin plot of `cholesterol` grouped by `continent` using the `gm2004` dataset. Restrict to three continents of your choice.

5. Save the interactive plot from Exercise 3 as an HTML file named `"gm_interactive.html"` using `htmlwidgets::saveWidget()`.
