## The Native Pipe Operator {#native-pipe}

In Section \@ref(filter-observations), we introduced the pipe operator `%>%` from the **magrittr** package (loaded automatically with the tidyverse). Starting from **R version 4.1.0** (released in 2021), R includes a built-in **native pipe operator** `|>` that works without loading any packages.

### Syntax Comparison

Both pipe operators take the result of the expression on the left and pass it as the **first argument** to the function on the right.

```{r, results=TRUE}
library(r02pro)
library(dplyr)

# Using the magrittr pipe
sahp %>%
  filter(house_style == "2Story") %>%
  summarize(avg_price = mean(sale_price))

# Using the native pipe
sahp |>
  filter(house_style == "2Story") |>
  summarize(avg_price = mean(sale_price))
```

As you can see, they produce exactly the same result.

### Key Differences

While the two pipe operators behave similarly in most cases, there are a few important differences:

1. **No placeholder**: The `%>%` pipe supports `.` as a placeholder to insert the left-hand side into a position other than the first argument. The native pipe `|>` does not support a placeholder in the same way. Instead, you can use an anonymous function.

```{r, results=TRUE}
# magrittr pipe with placeholder
c(3, 1, 2) %>% paste("Item", .)

# native pipe with anonymous function
c(3, 1, 2) |> (\(x) paste("Item", x))()
```

2. **No package needed**: The native pipe `|>` is built into R, so you don't need to load any package to use it.

3. **Speed**: The native pipe is slightly faster since it is implemented at the language level rather than as a function call.

### Which One Should You Use?

Both pipes are widely used in the R community. Here are some guidelines:

- If you are writing a **tidyverse-heavy** script and already loading **dplyr** or **magrittr**, either pipe works well.
- If you want to **minimize package dependencies**, use the native pipe `|>`.
- If you need the **placeholder** feature frequently, `%>%` may be more convenient.
- In this book, we primarily use `%>%` for consistency with existing examples, but you are encouraged to try `|>` in your own work.

### Exercises

1. Rewrite the following code using the native pipe `|>` instead of `%>%`:
```{r, eval = FALSE}
sahp %>%
  filter(oa_qual > 5) %>%
  select(sale_price, liv_area) %>%
  arrange(desc(sale_price)) %>%
  head(10)
```

2. Using the `gm2004` dataset, write a pipeline with the native pipe `|>` that filters for countries in `"Europe"`, selects `country` and `life_expectancy`, and arranges by `life_expectancy` in descending order.

3. Explain why the following code using the native pipe does not work, and rewrite it so that it does:
```{r, eval = FALSE}
c(1, 5, 3, 2, 4) |> sort() |> paste("Value:", .)
```
