## Hypothesis Testing {#hypothesis-testing}

In this section, we introduce the basics of hypothesis testing in R. Hypothesis testing is a fundamental tool in statistics that allows us to make inferences about population parameters based on sample data.

### The t-test

The **t-test** is used to test whether the mean of a population (or the difference between two population means) equals a specified value. R provides the `t.test()` function for this purpose.

#### One-sample t-test

Suppose we want to test whether the average sale price of houses in the `sahp` dataset is different from \$200,000 (i.e., 200 in the unit of thousands).

```{r, results=TRUE}
library(r02pro)
t.test(sahp$sale_price, mu = 200)
```

The output includes:

- **t**: the test statistic
- **df**: degrees of freedom
- **p-value**: the probability of observing a test statistic as extreme as the one computed, assuming the null hypothesis is true
- **95 percent confidence interval**: a range of plausible values for the true mean
- **sample estimates**: the sample mean

If the p-value is less than our significance level (commonly 0.05), we reject the null hypothesis.

#### Two-sample t-test

We can also compare the means of two groups. Let's test whether one-story and two-story houses have different average sale prices.

```{r, results=TRUE}
price_1story <- sahp$sale_price[sahp$house_style == "1Story"]
price_2story <- sahp$sale_price[sahp$house_style == "2Story"]
t.test(price_1story, price_2story)
```

By default, `t.test()` performs a **Welch two-sample t-test**, which does not assume equal variances. To assume equal variances, set `var.equal = TRUE`.

```{r, results=TRUE}
t.test(price_1story, price_2story, var.equal = TRUE)
```

#### Paired t-test

A **paired t-test** is used when observations come in pairs (e.g., before/after measurements). You can specify `paired = TRUE` in `t.test()`.

### The Chi-Squared Test

The **chi-squared test** is used to test the association between two categorical variables. R provides `chisq.test()` for this purpose.

Let's test whether kitchen quality (`kit_qual`) and house style (`house_style`) are independent in the `sahp` dataset.

```{r, results=TRUE}
# Create a contingency table
quality_table <- table(sahp$kit_qual, sahp$house_style)
quality_table

# Perform chi-squared test
chisq.test(quality_table)
```

A small p-value suggests that the two variables are not independent.

### The Correlation Test

To test whether the correlation between two numeric variables is significantly different from zero, use `cor.test()`.

```{r, results=TRUE}
cor.test(sahp$liv_area, sahp$sale_price)
```

The output shows the estimated correlation, the test statistic, and the p-value.

### Interpreting Results

When interpreting hypothesis test results, keep these points in mind:

1. A **small p-value** (typically < 0.05) provides evidence against the null hypothesis, but does not prove that the alternative is true.
2. **Statistical significance** does not imply **practical significance**. A very large sample can make tiny differences statistically significant.
3. The **confidence interval** provides more information than the p-value alone, as it shows the range of plausible values for the parameter.

### Exercises

1. Using the `gm2004` dataset from the **r02pro** package, perform a one-sample t-test to test whether the average `life_expectancy` across all countries is different from 70. What do you conclude at the 0.05 significance level?

2. Using the `gm2004` dataset, perform a two-sample t-test to compare the average `cholesterol` between `"male"` and `"female"` groups (the `gender` variable). Is there a statistically significant difference?

3. Create a contingency table of `continent` and `gender` from the `gm2004` dataset. Perform a chi-squared test to determine whether continent and gender are independent.

4. Test whether the correlation between `BMI` and `cholesterol` in the `gm2004` dataset is statistically significant. Report the estimated correlation and the p-value.

5. Using the `sahp` dataset, test whether houses with central air conditioning (`central_air == "Y"`) have a significantly higher average `sale_price` than those without. Use a two-sample t-test with `var.equal = FALSE`.
