## Linear Regression {#linear-regression}

In this section, we introduce **linear regression**, one of the most widely used statistical models. Linear regression allows us to model the relationship between a continuous response variable and one or more predictor variables.

### Simple Linear Regression

A simple linear regression models the relationship between a response variable $y$ and a single predictor variable $x$ as:
$$y = \beta_0 + \beta_1 x + \epsilon,$$
where $\beta_0$ is the intercept, $\beta_1$ is the slope, and $\epsilon$ is the error term.

In R, we use the `lm()` function (short for "linear model") to fit a regression. The syntax uses a **formula**: `y ~ x`.

```{r, results=TRUE}
library(r02pro)
model <- lm(sale_price ~ liv_area, data = sahp)
summary(model)
```

The `summary()` output includes:

- **Coefficients**: the estimated intercept ($\hat\beta_0$) and slope ($\hat\beta_1$), along with their standard errors, t-values, and p-values.
- **R-squared**: the proportion of variance in the response explained by the predictor. A value closer to 1 indicates a better fit.
- **F-statistic**: tests whether the overall model is statistically significant.

In this example, the slope for `liv_area` tells us the estimated change in `sale_price` for each additional unit increase in living area.

### Visualizing the Regression Line

You can easily overlay the regression line on a scatterplot using `geom_smooth()` in **ggplot2**.

```{r, results=TRUE}
library(ggplot2)
ggplot(sahp, aes(x = liv_area, y = sale_price)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(title = "Sale Price vs. Living Area",
       x = "Living Area (sq ft)",
       y = "Sale Price (thousands)")
```

The shaded region around the line represents the 95% confidence interval for the regression line.

### Multiple Linear Regression

You can include multiple predictors by adding them to the formula with `+`.

```{r, results=TRUE}
model_multi <- lm(sale_price ~ liv_area + lot_area + oa_qual, data = sahp)
summary(model_multi)
```

Each coefficient represents the estimated effect of that predictor while holding the other predictors constant.

### Extracting Model Information

R provides several useful functions to extract information from a fitted model.

```{r, results=TRUE}
# Fitted (predicted) values
head(fitted(model))

# Residuals (observed - fitted)
head(residuals(model))

# Coefficients
coef(model)

# Confidence intervals for coefficients
confint(model)
```

### Diagnostic Plots

Checking model assumptions is an important step in regression analysis. R provides built-in diagnostic plots.

```{r, results=TRUE, fig.width=7, fig.height=6}
par(mfrow = c(2, 2))
plot(model)
```

These four plots help you check:

1. **Residuals vs Fitted**: linearity and homoscedasticity (constant variance)
2. **Normal Q-Q**: normality of residuals
3. **Scale-Location**: another view of homoscedasticity
4. **Residuals vs Leverage**: influential observations

### Exercises

1. Using the `sahp` dataset, fit a simple linear regression of `sale_price` on `lot_area`. Interpret the estimated slope coefficient.

2. Using the `gm2004` dataset from the **r02pro** package, fit a simple linear regression of `life_expectancy` on `GDP_per_capita`. What is the R-squared value? What does it tell you?

3. Create a scatterplot of `life_expectancy` vs. `GDP_per_capita` from Exercise 2 and overlay the regression line using `geom_smooth(method = "lm")`.

4. Fit a multiple linear regression of `sale_price` on `liv_area`, `lot_area`, `oa_qual`, and `yr_built` using the `sahp` dataset. Which predictors are statistically significant at the 0.05 level?

5. For the model in Exercise 4, generate the four diagnostic plots using `plot()`. Do you see any violations of the model assumptions?
