## Error Handling and Debugging {#error-handling}

When writing R code, you will inevitably encounter errors, warnings, and unexpected behavior. In this section, you will learn how to handle these situations gracefully and debug your code effectively.

### Errors, Warnings, and Messages

R has three types of signals that communicate different levels of severity:

- **Errors** (produced by `stop()`): The code cannot continue. Execution halts immediately.
- **Warnings** (produced by `warning()`): Something unexpected happened, but the code can still continue.
- **Messages** (produced by `message()`): Informational output that does not indicate a problem.

```{r, results=TRUE, error=TRUE}
log(-1)        # produces a warning
log("abc")     # produces an error
message("This is an informational message.")
```

### Generating Custom Errors and Warnings

When writing your own functions, you should signal problems clearly to the user.

```{r, results=TRUE, error=TRUE}
safe_divide <- function(x, y) {
  if (y == 0) {
    stop("Division by zero is not allowed.")
  }
  if (x == 0) {
    warning("Numerator is zero; result will be zero.")
  }
  x / y
}
safe_divide(10, 2)
safe_divide(0, 5)
safe_divide(10, 0)
```

### Handling Errors with `try()` and `tryCatch()`

#### `try()`: Keep Going After an Error

The `try()` function wraps an expression so that if it produces an error, execution continues instead of stopping.

```{r, results=TRUE}
result <- try(log("abc"), silent = TRUE)
class(result)
```

When `silent = TRUE`, the error message is suppressed. You can check whether an error occurred using `inherits(result, "try-error")`.

```{r, results=TRUE}
if (inherits(result, "try-error")) {
  cat("An error occurred, using a default value instead.\n")
  result <- NA
}
result
```

#### `tryCatch()`: Fine-Grained Error Handling

The `tryCatch()` function provides more control by letting you specify handlers for different types of conditions.

```{r, results=TRUE}
safe_log <- function(x) {
  tryCatch(
    log(x),
    warning = function(w) {
      message("Warning caught: ", conditionMessage(w))
      NA
    },
    error = function(e) {
      message("Error caught: ", conditionMessage(e))
      NA
    }
  )
}
safe_log(10)
safe_log(-1)
safe_log("abc")
```

A practical use case is applying a function to a list of inputs where some might cause errors:

```{r, results=TRUE}
inputs <- list(4, -1, "hello", 100)
results <- sapply(inputs, safe_log)
results
```

### Debugging Tools

When your code produces unexpected results but no error, you need to investigate. R provides several debugging tools.

#### `traceback()`: Find Where an Error Occurred

After an error, calling `traceback()` shows the sequence of function calls that led to the error.

```{r, eval=FALSE}
f <- function(x) g(x)
g <- function(x) h(x)
h <- function(x) log(x)
f("abc")
traceback()
```

#### `browser()`: Pause and Inspect

Inserting `browser()` inside a function pauses execution at that point and opens an interactive debugging session where you can inspect variables.

```{r, eval=FALSE}
my_function <- function(x) {
  y <- x^2
  browser()  # execution pauses here
  z <- y + 1
  z
}
my_function(5)
```

In the debugging session, you can type variable names to see their values, `n` to step to the next line, `c` to continue, or `Q` to quit.

#### `debug()` and `undebug()`

You can also set a function to be debugged without modifying its source code:

```{r, eval=FALSE}
debug(my_function)
my_function(5)      # will enter browser mode automatically
undebug(my_function) # stop debugging
```

### Exercises

1. Write a function `safe_sqrt(x)` that uses `tryCatch()` to return `NA` when the input is negative (instead of producing a warning) and when the input is not numeric (instead of producing an error). Test it with `safe_sqrt(9)`, `safe_sqrt(-4)`, and `safe_sqrt("hello")`.

2. Given a list of file paths `files <- c("data/file1.csv", "data/file2.csv", "data/nonexistent.csv")`, write code using `try()` inside a `for` loop that attempts to read each file with `read.csv()` and prints a message for files that cannot be read. (You do not need actual files; just observe the error handling.)

3. Write a function `divide_all(x, y)` that divides each element of vector `x` by the corresponding element of vector `y`. Use `tryCatch()` to handle cases where `y` contains zeros, returning `Inf` for those positions instead of an error.

4. Use `browser()` inside a simple function to practice stepping through code. Write a function that computes the cumulative sum of a vector using a `for` loop, and use `browser()` to inspect the intermediate values.
