## The Apply Family and Purrr {#apply-purrr}

In Chapters \@ref(control-statements) and \@ref(functions), you learned how to write loops and define functions. A natural next question is: how can we **apply** a function to every element of a vector or list without writing an explicit `for` loop? R provides two powerful approaches: the base R **apply family** of functions and the **purrr** package from the tidyverse.

### The Apply Family

The base R apply family includes several functions that apply a given function to elements of a data structure. The most commonly used ones are `lapply()`, `sapply()`, and `vapply()`.

#### `lapply()`: Apply and Return a List

The function `lapply(X, FUN)` applies the function `FUN` to each element of `X` and always returns a **list**.

```{r, results=TRUE}
my_vec <- c(1, 4, 9, 16, 25)
lapply(my_vec, sqrt)
```

You can also use `lapply()` on a list.

```{r, results=TRUE}
my_list <- list(a = 1:5, b = 6:10, c = 11:15)
lapply(my_list, mean)
```

#### `sapply()`: A Simplified `lapply()`

The function `sapply()` works like `lapply()`, but tries to **simplify** the result into a vector or matrix when possible.

```{r, results=TRUE}
sapply(my_vec, sqrt)
sapply(my_list, mean)
```

Notice that `sapply()` returns a named numeric vector instead of a list, which is often more convenient.

#### `vapply()`: A Safer `sapply()`

The function `vapply()` is similar to `sapply()`, but requires you to specify the expected return type. This makes it safer for programming because it will throw an error if the output doesn't match the expected type.

```{r, results=TRUE}
vapply(my_list, mean, numeric(1))
```

Here, `numeric(1)` tells R that each result should be a single numeric value.

#### `apply()`: For Matrices and Arrays

The function `apply(X, MARGIN, FUN)` applies a function over rows (`MARGIN = 1`) or columns (`MARGIN = 2`) of a matrix.

```{r, results=TRUE}
my_mat <- matrix(1:12, nrow = 3)
my_mat
apply(my_mat, 1, sum)  # row sums
apply(my_mat, 2, sum)  # column sums
```

### Introduction to Purrr

The **purrr** package provides a more consistent and flexible alternative to the apply family. Its core function is `map()`, which always returns a list.

```{r, results=TRUE}
library(purrr)
map(my_list, mean)
```

#### Typed Map Functions

One advantage of **purrr** is that it offers typed variants of `map()` that specify the output type:

- `map_dbl()`: returns a double (numeric) vector
- `map_chr()`: returns a character vector
- `map_lgl()`: returns a logical vector
- `map_int()`: returns an integer vector

```{r, results=TRUE}
map_dbl(my_list, mean)
map_chr(my_list, ~ paste("Mean:", round(mean(.x), 1)))
```

In the second example, we used a **formula notation** (`~ ...`) to define an anonymous function inline. The `.x` refers to the current element being processed.

#### `map_df()`: Return a Data Frame

When each element returns a data frame (or a named vector), you can use `map_df()` to combine the results into a single data frame.

```{r, results=TRUE}
library(r02pro)
map_df(my_list, ~ tibble(mean = mean(.x), sd = sd(.x)), .id = "group")
```

### Comparison: For Loops vs. Apply vs. Purrr

Let's compare the three approaches for computing the mean of each column in a data frame.

```{r, results=TRUE}
library(r02pro)
sahp_numeric <- sahp[, c("sale_price", "liv_area", "lot_area")]

# Using a for loop
result_loop <- numeric(ncol(sahp_numeric))
for (i in seq_along(sahp_numeric)) {
  result_loop[i] <- mean(sahp_numeric[[i]], na.rm = TRUE)
}
names(result_loop) <- names(sahp_numeric)
result_loop

# Using sapply
sapply(sahp_numeric, mean, na.rm = TRUE)

# Using purrr
map_dbl(sahp_numeric, mean, na.rm = TRUE)
```

All three approaches produce the same result. The `sapply()` and `map_dbl()` versions are more concise and clearly express the intent: apply a function to each element.

**When to use which:**

- Use `for` loops when the iteration has side effects or when each step depends on the previous one.
- Use `sapply()`/`vapply()` for quick, base-R-only scripts.
- Use `map()` and its variants for tidyverse workflows, especially when chaining with the pipe operator.

### Exercises

1. Given the list `my_data <- list(x = c(1, 3, 5, NA), y = c(2, 4, 6, 8), z = c(10, 20, 30))`, use `sapply()` to compute the mean of each element, handling `NA` values with `na.rm = TRUE`.

2. Repeat Exercise 1 using `map_dbl()` from the **purrr** package.

3. Create a matrix with 4 rows and 5 columns filled with random numbers (use `matrix(rnorm(20), nrow = 4)`). Use `apply()` to compute the standard deviation of each column.

4. Using the `sahp` dataset, select the numeric columns `sale_price`, `liv_area`, `lot_area`, and `oa_qual`. Use `map_df()` to create a summary data frame with the mean, median, and standard deviation of each variable.

5. Write a function `scale_vec(x)` that scales a numeric vector to have mean 0 and standard deviation 1. Use `lapply()` to apply this function to each numeric column of the `sahp` subset from Exercise 4.
